Modeling Second-Language Learning from a Psychological Perspective

### 2018-03-26 doc/naaclhlt2018-latex/paper.tex
add lesions graph and description (no interpretation of results yet)

\documentclass[11pt,a4paper]{article}
\usepackage[hyperref]{naaclhlt2018}
\usepackage{times}
\usepackage{latexsym}
\usepackage{soul}

@@

To account for the effects of surrounding words on the difficulty of an
instance, we created several features related to the instance word's context in
the exercise. These included the token of the previous word, the next word, and
the instance word's root in the parse tree, all stored in single columns as with
the instance token itself. We also included the part of speech of each of these
context words as additional features. When there was no previous word, next word, or parse
root word, a special {\tt None} token or {\tt None} part of speech was used.

@@

@@

\begin{table}[t!]
\small
\begin{center}
\begin{tabular}{|l|llll|}
  % \hline   & \multicolumn{3}{l}{\bf Value} & \\
  \hline \bf Parameter & {\tt fr\_en} & {\tt en\_es} & {\tt es\_en} & {\tt all} \\ \hline
  num\_leaves & 256 & 512 & 512 & 1024 \\
  learning\_rate & .05 & .05 & .05 & .05 \\
  min\_data\_in\_leaf & 100 & 100 & 100 & 100 \\
  num\_boost\_rounds & 750 & 650 & 600 & 750 \\
  cat\_smooth & 200 & 200 & 200& 200 \\
  feature\_fraction & .7 & .7 & .7 & .7 \\
  max\_cat\_threshold & 32 & 32 & 32& 64 \\

\hline
\end{tabular}
\end{center}
\caption{\label{lightgbm-params} Parameters of final LightGBM models. See
  LightGBM documentation for more information; all other parameters were left at
their default values.}
\end{table}

@@

The AUROC of our final predictions was $.8585$ on {\tt en\_es}, $.8350$ on {\tt
  es\_en}, and $.8540$ on {\tt fr\_en}. We did not attempt to optimize the model's
F1 score, but the model's F1 score could likely be improved (at the cost of
increased log loss) by finding the
rescaling of the ``dev'' predicted probabilities that maximized the F1 score at
the 0.5 threshold, and applying this rescaling to the ``test'' predicted probabilities.

@@

>>>>

\documentclass[11pt,a4paper]{article}
\usepackage[hyperref]{naaclhlt2018}
\usepackage{times}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{soul}

@@

To account for the effects of surrounding words on the difficulty of an
instance, we created several features related to the instance word's context in
the exercise. These included the token of the previous word, the next word, and
the instance word's root in the dependency tree, all stored in single columns as with
the instance token itself. We also included the part of speech of each of these
context words as additional features. When there was no previous word, next word, or dependency-tree
root word, a special {\tt None} token or {\tt None} part of speech was used.

@@

\begin{table}[t!]
\small
\begin{center}
\begin{tabular}{|l|llll|}
  % \hline   & \multicolumn{3}{l}{\bf Value} & \\
  \hline \bf Parameter & {\tt fr\_en} & {\tt en\_es} & {\tt es\_en} & {\tt all} \\ \hline
  num\_leaves & 256 & 512 & 512 & 1024 \\
  learning\_rate & .05 & .05 & .05 & .05 \\
  min\_data\_in\_leaf & 100 & 100 & 100 & 100 \\
  num\_boost\_rounds & 750 & 650 & 600 & 750 \\
  cat\_smooth & 200 & 200 & 200& 200 \\
  feature\_fraction & .7 & .7 & .7 & .7 \\
  max\_cat\_threshold & 32 & 32 & 32& 64 \\
\hline
\end{tabular}
\end{center}
\caption{\label{lightgbm-params} Parameters of final LightGBM models. See
  LightGBM documentation for more information; all other parameters were left at
their default values.}
\end{table}

@@

@@

The AUROC of our final predictions was $.8585$ on {\tt en\_es}, $.8350$ on {\tt
  es\_en}, and $.8540$ on {\tt fr\_en}. We did not attempt to optimize the model's
F1 score, but the F1 score could likely be improved (at the cost of
increased log loss) by finding the
rescaling of the ``dev'' predicted probabilities that maximized the F1 score at
the 0.5 threshold, and applying this rescaling to the ``test'' predicted probabilities.

@@

\begin{figure*}[htp]
% \centering
\includegraphics[width=\textwidth]{lesions.pdf}
\caption{Performance on {\tt dev} of models trained on all {\tt train} data,
  with different groups of lesioned features. See main text for description of
  lesion groups}
\label{fig:lesions}
\end{figure*}

To better understand which features or groups of features were most important to
our model's predictions, we conducted a set of experiments in which we lesioned
(i.e., removed) a group of features and re-trained the model on the {\tt train}
set, evaluating performance on the {\tt dev} set. For simplicity, we ran each of the
lesioned models on all language data and report the average
performance. We did not run individual-language models as we did for our primary model.

The results of the lesion experiments are shown in Figure~\ref{fig:lesions}. The
models are as follows.
\begin{itemize}
  \item \emph{none}: all features are included.  
  \item \emph{user}: all user-level features, including the user ID and other
    calculated features like entropy and measures of exercise bursts, are
    removed.
  \item \emph{userid} and \emph{user other}: only user ID or only the calculated
    features user features, respectively, are removed.
  \item \emph{word}: Token and token root IDs; previous, next, and dependency-tree
    root word IDs; and morphological, part of speech, and dependency tree
    features are removed.
  \item \emph{word id} and \emph{word other}: only word IDs or only other
    word features, respectively, are removed.
  \item \emph{neighbors}: Both word IDs and other word features are removed, but
    only for the previous, next, and dependency-tree root words. Information
    about the present word is maintained.
  \item \emph{external}: External information about the word, including corpus
    frequency, Levenshtein distance from translation, and age of acquisition,
    are removed.
  \item \emph{temporal}: Temporal information, including number and timing of
    past encounters with the word and error tracking information, is removed.
\end{itemize}
